---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

```{r}
library(dplyr)
library(tidyr)
library(readr)

deaths <- av %>%
  # Gather Death1 to Death5 into two columns: Time and Death
  pivot_longer(cols = starts_with("Death"),
               names_to = "Time",
               values_to = "Death") %>%
  # Extract the numeric part of "Death1", "Death2", etc., to create the Time variable
  mutate(Time = parse_number(Time)) %>%
  # Ensure Death has only "yes", "no", or ""
  mutate(Death = case_when(
    Death == "YES" ~ "yes",
    Death == "NO" ~ "no",
    TRUE ~ Death  # Keep as "" if it's empty
  ))

```


Similarly, deal with the returns of characters.

```{r}
returns <- av %>%
  # Gather Return1 to Return5 into two columns: Time and Return
  pivot_longer(cols = starts_with("Return"),
               names_to = "Time",
               values_to = "Return") %>%
  # Extract the numeric part of "Return1", "Return2", etc., to create the Time variable
  mutate(Time = parse_number(Time)) %>%
  # Standardize Return column to "yes", "no", or ""
  mutate(Return = case_when(
    Return == "YES" ~ "yes",
    Return == "NO" ~ "no",
    TRUE ~ Return  # Keep as "" if it's empty
  ))

summary(returns)
```



Based on these datasets calculate the average number of deaths an Avenger suffers. 

```{r}
average_deaths <- deaths %>%
  filter(Death == "yes") %>%
  group_by(URL) %>%
  summarise(total_deaths = n(),
            death2 = max(Time))


mean(average_deaths$total_deaths)
sum(average_deaths$total_deaths)
```



## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement

### Include your answer

```{r}
# Load necessary libraries
library(dplyr)

# Define parameters
years_in_operation <- 53
months_in_operation <- years_in_operation * 12

# Calculate total deaths
total_deaths <- deaths %>%
  filter(Death == "yes") %>%
  summarise(total_deaths = n()) %>%
  pull(total_deaths)

# Calculate death frequency
death_frequency <- months_in_operation / total_deaths

# Display the frequency of deaths
death_frequency

```


Include at least one sentence discussing the result of your fact-checking endeavor.

I found that it is roughly correct, every 7.146 months on average an avenger dies

Upload your changes to the repository. Discuss and refine answers as a team.
